Study of the hadronization of charged pions

### 2016-10-20 chap5.tex
Changes from Lamiaa

This chapter treats the analysis of CLAS data taken for both 
experiments E-02-104 \cite{Brooks:2002aa} and E-02-110 \cite{Hafidi:2002aa}.
The run period, named {\it eg2}, was composed of three phases 
labeled {\it a}, {\it b} and {\it c}. The analysis is performed on the data 
collected in the beginning of 2004 during the third phase ({\it eg2c}), for 
which the beam energy was 5.014~GeV\footnote{The other phases of the run gave only small 
amounts of data at beam energies of 4 and 4.7~GeV, Thus, we decided not to use them 
for this analysis.}.

As the two experiments, running simultaneously, aimed at comparing deuterium 
with heavier nuclei, it was decided to use a double target system 
\cite{Hakobyan:2008zz}. The first target was filled with liquid deuterium and 
the second was a solid target. The latter was made of either carbon, aluminum, 
iron, tin or lead and was changed remotely (the system is shown in figure 
\ref{fig:phototarget}). The two targets were separated by only 4~cm in order to reduce
the acceptance differences between them. The 
advantage of having two targets simulataneously in the beam line
is that several systematic effects related to the beam 
and detector properties will cancel in the nuclear ratio.

\begin{figure}[htbp]
\centering
\includegraphics[width=7cm] {chap5-fig/PhTar.jpg} 
\caption {Picture of the target system of the {\it eg2} run. The cryogenic target is 
in the back, wrapped with aluminum foils. The solid targets are held by mechanical arms
that were remotely controled. In the picture, the top solid target is in the beam line.}
\label{fig:phototarget}
\end{figure}

The analysis of the experiment E-02-110, focusing on the study of color 
transparency effects, has been approved and published~\cite{ElFassi:2008}. As this 
analysis went through a careful review by the CLAS collaboration, we use
similar analysis methods as much as possible. In particular, the electron selection 
presented here is very similar to theirs; the main difference being a new 
target determination method. The pion identification methods are however tightened
in this analysis as we have less constraints on the kinematics to exclude possible
contaminations.

In the section \ref{sec:pid}, we identify the following particles e$^-$, 
$\pi^-$ and $\pi^+$ using a series of cuts on the various 
detector outputs. For pions, we use the signal in the drift chambers (DC) and 
in the scintillator counters (SC), for which we require a positive status (global and DC) 
from the reconstruction code. For electron identification, we use signals from 
all detector parts DC, Cherenkov Counters (CC), SC and Electromagnetic 
Calorimeter (EC).

Once particles are identified, we can extract the observables (multiplicity 
ratios and transverse momentum broadening). The method is presented in the 
section \ref{sec:obs}, the corrections to these observables are presented in the 
section \ref{sec:corrections}. The evaluation of 
the systematic errors is presented in the section \ref{sec:TotSys}.
Finally, the analysis' final results are presented and discussed in the next chapter.

@@

First, we apply a fiducial cut on the EC to remove the electrons detected on 
the edge of the calorimeter ($U_{EC}>40$~cm, $V_{EC}<360$~cm and 
$W_{EC}<395$~cm in the calorimeter's coordinates). These are problematic hits 
because the generated electromagnetic shower might be partly outside the 
detector, leading to a wrong measurement of the energy deposited.

To reject pions, we apply cuts on the energy deposited
in the EC using the measurements in the inner ($E_{in}$) and the 
outer part ($E_{out})$ of the calorimeter:

@@

where $\mu = 0.271$ is the mean of the fraction of the energy deposited in the 
calorimeter by electrons, the $a$ parameter is set at 0.5 and the parameter 
$b$ is a function of momentum given in table \ref{tab:ecoutin-par}. The 
adjustment of $b$ is motivated by the non-linear dependence of the energy 
deposited as a function of the particle's momentum (see figure \ref{eleEC}). As pions are 
minimum ionizing particles, they are expected to lose a constant energy in 
the inner part of the calorimeter (around 30~MeV), regardless of their 
momentum. Therefore, by requesting more than 60~MeV to be deposited, we 
efficiently cut the pion contamination (see figure~\ref{eleECi}).

@@

\begin{figure}[p]
\centering
\includegraphics[width=14cm] {chap5-fig/fig02.jpg} 
\caption {Energy deposited in the inner part of the EC ($E_{in}$) as a function of the 
energy deposited in the outer part ($E_{out}$), both divided by the momentum of the 
particle. Each panel is for a different momentum range, the red lines 
illustrate the cuts from equation \ref{einout}.}
\label{eleEC}
\end{figure}

\begin{figure}[tbp]
\centering
\includegraphics[width=9cm] {chap5-fig/fig03.png} 
\caption {Energy deposited in the inner part of the EC as a function of the 
energy in the outer part. The red line illustrates the cut applied for 
electron selection.}
\label{eleECi}
\end{figure}

In the CC, the mean number of photo-electrons\footnote{Photo-electrons are 
electrons produced in the front window of the Photo-Multiplier Tube (PMT) by 
single photons.} from a high energy electron is expected to be around 10.
However, hadrons can generate noise due to $\delta$ electrons produced in 
the materials of the detector. This signal is expected to be around 
one to two photo-electrons, hence to remove it we keep only tracks with more than 
2.5~photo-electrons (figure \ref{delta}).

\begin{figure}[tbp]
\centering
\includegraphics[width=8cm] {chap5-fig/fig04.png} 
\caption {Number of photo-electrons ($\times 10$) per tracks, the red line 
shows the cut to remove pion contamination.}
\label{delta}
\end{figure}

To select electrons, we also require that the particles are negatively charged, 
which is determined based on the bending direction of the track in the drift chamber.
Positively charged particles verifying electrons' cuts are identified as 
positrons. In the rest of the 
analysis we use only events with a single electron and no positron to avoid any 
confusion between the scattered electron and electrons originating from hadronic decays or 
a photon conversion.

@@

To identify negatively charged pions, we select negative tracks, which are not 
identified as electron. Pions are detected in the angular range from $\sim$10 to 
$\sim$140 degrees using the DC and SC only.

The identification consists mostly of a time of flight (TOF) test. We define 
the relativistic particles' velocity difference as
$\Delta \beta = \beta_{measured} - {p \over \sqrt{p^2 + m_\pi^2}}$ and request 
$\Delta \beta$ to be zero within $\pm 0.03$. Figure \ref{PionTOF} illustrates 
the effect of this cut. We notice that there is not much negative kaons or 
anti-protons. Therefore, the contamination from these should be small (see 
section \ref{SysId} for more detailed analysis).

\begin{figure}[tbp]
\centering
\includegraphics[width=8cm] {chap5-fig/fig05.png} 
\caption {$\Delta \beta$ as a 
function of momentum (GeV/c). Only negative particles are plotted. The red 
lines are the cuts applied to select negative pions.}
\label{PionTOF}
\end{figure}

In principle, pion identification could be improved using the Cherenkov counter 
for momentum higher than 2.5~GeV/c. But, the low efficiency observed (figure 
\ref{PionCC}), especially at momentum close to the threshold ($\sim$25\% at 
2.5~GeV/c and $\sim$50\% at 3~GeV/c), makes its use less compelling. 
Moreover, as only a very small amount of $K^-$ and no $\bar p$ are present on 
the figure \ref{PionTOF}, we decided to not use the CC for the pion identification.

\begin{figure}[tbp]
\centering
\includegraphics[width=8cm] {chap5-fig/fig06.png} 
\caption {Histograms of momentum (GeV/c) for $\pi^-$. In black are all 
identified $\pi^-$, in red pions that also fire in the Cherenkov counter and 
in blue those that fire in the Cherenkov counter with more than 
1 photo-electrons.}
\label{PionCC}
\end{figure}

@@

The identification of positively charged pions is similar to that of negative 
pions. However, the time of flight plot is significantly more busy (figure 
\ref{PipTOF}), showing a significant contamination from $K^+$ and protons at 
high momentum. As the CC is not efficient enough for hadron separation, the 
numerous kaons and protons should be removed by a tighter TOF cut:

@@

These cuts permit to minimize the kaon contamination, below 2.5~GeV/c, and the 
proton contamination for all momentum. The kaon contribution cannot be avoided, however 
it should remain a small contribution ($\sim$ 3\% according to simulation) 
with only a small effect on the final results (see section \ref{SysId} 
for details). Because protons could lead to even more contamination than kaons, 
a stricter cut is used at high momentum. This cut is also justified by 
HERMES data \cite{Airapetian:2007vu}, which show a very 
different behavior of the protons compared to the other hadrons.

\begin{figure}[tbp]
\centering
\includegraphics[width=8cm] {chap5-fig/pip_data.png} 
\caption {$\Delta \beta$ as a 
function of momentum (GeV/c). Only positive particles are shown, the red 
lines are the cuts applied to select positive pions. The blue lines indicate 
the theoretical positions for other particles, we plotted from top to 
bottom the lines for positrons, muons, kaons and protons.}
\label{PipTOF}
\end{figure}

@@

To differentiate the two targets and remove the background, we need to 
determine the origin of the particles. It appears that the different sectors 
are shifted in $z$ (figure \ref{vertex}), because of a small misalignment of 
the beam with the detector. To correct this problem, we apply the modification shown in table 
\ref{tab:vertex} for the vertex determination. Those values were determined by 
fitting the solid target position viewed by each sector.

\begin{table}[p]
  \centering
  \begin{tabular}{@{} cc @{}}
    \hline
    Sector & Shift (cm) \\ 
    \hline
    1 & + 0.1 \\ 
    2 & - 0.4 \\ 
    3 & - 0.6 \\ 
    4 & - 0.1 \\ 
    5 & + 0.4 \\ 
    6 & + 0.6 \\ 
    \hline
  \end{tabular}
  \caption{Values used to correct the vertex information for all sectors}
  \label{tab:vertex}
\end{table}

The position of the targets may also vary from one run to another, but 
this does not happen too often. Indeed, except for aluminum, the targets remain in 
the same position within one or two millimeters. The positions, in CLAS 
coordinates, are given in table \ref{tab:targets}.

\begin{table}[p]
  \centering
  \begin{tabular}{|c|c|c|c|c|c|c|}
    \hline
    Target & Carbon & Al (1) & Al (2) & Iron   & Tin    & Lead   \\ 
    \hline \hline
    Liquid & -30.1  & n/a    & n/a    & -30.2  & -30.1  & -30.1  \\ 
    Solid  & -24.7  & -25.0  & -23.8  & -24.9  & -23.8  & -24.9  \\
    \hline
  \end{tabular}
  \caption{Measured mean vertex position of targets, relative to the center of 
           CLAS (in cm). We note that aluminum data are separated in two sets because the 
           target position has brutally changed, moreover
           during these runs the deuterium target was empty.}
  \label{tab:targets}
\end{table}

\begin{figure}[p]
\centering
\includegraphics[width=10cm] {chap5-fig/Vertex_el_data.png}
\caption {In each sector, the reconstructed vertex of electrons along the beam
direction relatively to the center of CLAS. The red lines show the cuts 
to select the targets.}
\label{vertex}
\end{figure}

The detected electrons are associated with the solid target if their vertex 
position is less than 1.5~cm ($\sim$3~$\sigma$) from the value of table 
\ref{tab:targets}. For the liquid target, the cut is larger, 2 cm, in order to 
account for the size of the target (see figure \ref{vertex}). The vertex of 
the pions is checked against the electron one and we request that 
$| Vz^{e^-} - Vz^{\pi} | < 3$ cm (see figure \ref{fig:dvzpi}).

\begin{figure}[tbp]
\centering
\includegraphics[width=8cm] {chap5-fig/Vertex_pi_data.png}
\caption {Distance along the beam axis between the electrons and pions 
vertexes. $\pi^+$ are plotted in the top panel and the $\pi^-$ in the bottom.
The red lines show the cuts applied to select pions.}
\label{fig:dvzpi}
\end{figure}

@@

To check the quality of the runs\footnote{Runs correspond roughly to two hours
of data, they can be smaller in case of problems.}, we looked at the ratio of the number of 
scattered electrons, between the liquid and solid targets. If the beam is 
hitting other materials than the targets or if the detector is not working 
properly this ratio can be off and indicates a problematic run. The figure 
\ref{DataQ} shows the values obtained; we fit the 
mean value for all runs of each target and eliminate runs away by more than 
5$\sigma$. We note that the ratios are coherent with target thicknesses given 
in \cite{Hakobyan:2008zz} except for carbon. Because of this, the density of 
the carbon target was remeasured recently and found to be coherent with the 
data\footnote{i.e. (1.747+/-0.0007)g/cm$^3$ instead of 2.235 g/cm$^3$}.

\begin{figure}[tbp]
\centering
\includegraphics[width=15cm] {chap5-fig/TargetElRatio.png}
\caption {Ratio of the number of electrons scattered by the targets (solid over deuterium) 
for each run.}
\label{DataQ}
\end{figure}

@@

Since we are interested in deep inelastic scattering, we use the following 
cuts: $Q^2 > 1$~GeV$^2$/c$^2$ and $W > 2$~GeV. We also apply a cut on the 
fraction of energy transferred $y = \nu/E_{beam} < 0.85$.
This corresponds to removing the low energy electrons because they are
significantly affected by radiative effects and sat the limit of our trigger 
threshold.

@@

Once the identification is done, the calculation of the observables is 
straight forward with statistical errors given by:
\begin{equation}
{\delta \left ( R_A^h \right ) \over R_A^h} = \sqrt{ 1 /N_A^h + 1/ N_A^e +1/N_D^h + 1/N_D^e}
\end{equation}
and
\begin{equation}
\left ( \delta \left ( \Delta \langle P_\perp^2 \rangle \right ) \right )^2 = 
   \left ({\langle P_\perp^4 \rangle - \langle P_\perp^2 \rangle ^2}\right )_A / N_A^h
 + \left ({\langle P_\perp^4 \rangle - \langle P_\perp^2 \rangle ^2}\right )_D / N_D^h.
\end{equation}

However, the implementation of acceptance and radiative corrections is done through weights
given to each events depending on their kinematics. The multiplicity ratio then becomes
\begin{equation}
R_A^h (Q^2,\nu,z_h,P_\perp^2) = {{\sum \omega_A^h (Q^2,\nu,z_h,P_\perp^2) / \sum \omega_A^e (Q^2,\nu)} 
                       \over {\sum \omega_D^h (Q^2,\nu,z_h,P_\perp^2) / \sum \omega_D^e (Q^2,\nu)}},
\end{equation}
with $\omega$ the weight associated with the event and the sums running over all measured 
particles. The expression for the transverse momentum broadening remains
\begin{equation}
\Delta \langle P_\perp^2 \rangle = \langle P_\perp^2 \rangle_A - \langle P_\perp^2 \rangle_D,
\end{equation}
but with
\begin{equation}
\langle P_\perp^2 \rangle = {\sum \left ( P_\perp^2 \times \omega (Q^2,\nu,z_h,P_\perp^2) \right ) \over \sum \omega (Q^2,\nu,z_h,P_\perp^2)}.
\end{equation}

Thus, new expressions are also needed for the statistical error uncertainties: 
\begin{equation}
{\delta \left ( R_A^h \right ) \over R_A^h} = 
      \sqrt{ \left ( {\sum {\omega^h_A}^2 \over \left (\sum \omega^h_A \right )^2} \right ) 
           + \left ( {\sum {\omega^e_A}^2 \over \left (\sum \omega^e_A \right )^2} \right ) 
           + \left ( {\sum {\omega^h_D}^2 \over \left (\sum \omega^h_D \right )^2} \right ) 
           + \left ( {\sum {\omega^e_D}^2 \over \left (\sum \omega^e_D \right )^2} \right ) }
\end{equation}
 and 
\begin{equation}
\begin{split}
\left ( \delta \left ( \Delta \langle P_\perp^2 \rangle \right ) \right )^2 = 
   \left ({{\sum \omega^h_A P_\perp^4 }\over{\sum \omega^h_A}} - \left ({\sum \omega^h_A P_\perp^2 }\over{\sum \omega^h_A}\right )^2 \right ) 
         \times \left ( {\sum \left ( {\omega^h_A}^2 \right ) \over \left ( \sum \omega^h_A \right ) ^2 }\right ) \\
 + \left ({{\sum \omega^h_D P_\perp^4 }\over{\sum \omega^h_D}} - \left ({\sum \omega^h_D P_\perp^2 }\over{\sum \omega^h_D}\right )^2 \right ) 
	 \times \left ( {\sum \left ( {\omega^h_D}^2 \right ) \over \left ( \sum \omega^h_D \right ) ^2 }\right ).
\end{split}
\end{equation}

@@

To provide a first idea on data quality, we present in figure~\ref{fig:prelim} 
few preliminary results, before the application 
of any corrections. These preliminary results will also be used to 
illustrate the effects of the corrections discussed below.

\begin{figure}[htb]
\centering
\includegraphics[width=7.4cm] {chap5-fig/a_RvA.png} 
\includegraphics[width=7.4cm] {chap5-fig/a_RvZ.png} 
\includegraphics[width=7.4cm] {chap5-fig/a_PvA.png} 
\includegraphics[width=7.4cm] {chap5-fig/a_PvNu.png} 
\caption {Results for multiplicity ratios (top) and transverse momentum 
broadening (bottom) without correction.}
\label{fig:prelim}
\end{figure}

@@

The acceptance correction consists of applying weights to the experimental data
events by events to correct for detection inefficiencies of the detectors.
Incidentally, it also corrects for small other issues of detection,
such as misidentification and scattering on detectors materials. The quality 
of the correction depends on the ability of the simulation to reproduce the 
experiment, the statistics accumulated and the size of interfering effects, such 
as bin migration. For this correction, we applied the method from the 
approved and published analysis from the same eg2 data~\cite{ElFassi:2008}.

@@

To correct for acceptance effects, we simulated a total of a 100 million events 
per target ($^2$H, C, Fe and Pb) using the PYTHIA \cite{Sjostrand:2006za} 
event generator slightly modified to include Fermi motion effects. The 
generated events are processed by the CLAS software (GSIM, GPP and user\_ana) 
to simulate the detector and the reconstruction process similarly to the one for 
experimental data.
% TODO GPP factors

>>>>

This chapter treats the analysis of CLAS data taken for both 
experiments E-02-104 \cite{Brooks:2002aa} and E-02-110 \cite{Hafidi:2002aa}.
The run period, named {\it eg2}, was composed of three phases 
labeled {\it a}, {\it b} and {\it c}. The analysis is performed on the data 
collected in the beginning of 2004 during the third phase ({\it eg2c}), for 
which the beam energy was 5.014~GeV\footnote{The other phases of the run gave only small amounts of data at beam energies of 4 and 4.7~GeV. Thus, we decided to not use them in this analysis.}.

As the two married experiments aimed at comparing deuterium with heavy nuclei, it was decided to use a double target system~\cite{Hakobyan:2008zz}. The first target was filled with a liquid deuterium, and the second was a solid target. The latter was made of either carbon, aluminum, iron, tin or lead, and was changed remotely (see figure~\ref{fig:phototarget}). The two targets were separated by only 4~cm in order to minimize the difference on CLAS acceptance between them. The advantage of having two targets mounted simultaneously on the beam line is that several systematic effects related to the beam and detector's properties will cancel in the nuclear ratio.

\begin{figure}[htbp]
\centering
\includegraphics[width=7cm] {chap5-fig/PhTar.jpg} 
\caption {Picture of the {\it eg2} target system. The cryogenic target is 
in the back wrapped with aluminum foils, while solid targets are held by mechanical arms that were controlled remotely. In the picture, the top solid target is in the beam line.}
\label{fig:phototarget}
\end{figure}

The analysis of the experiment E-02-110, focusing on the study of color 
transparency effects, has been approved and published~\cite{ElFassi:2008}. As this 
analysis went through a careful review by the CLAS collaboration, we have adapted similar analysis methods as much as possible. In particular, the electron selection 
presented here is very similar to theirs; the main difference being a new 
target determination method. The pion identification cuts are, however, tightened
in this analysis as we have less constraints on the kinematics to exclude any possible pions' contamination.

In the section \ref{sec:pid}, we identify the following particles, e$^-$, 
$\pi^-$ and $\pi^+$ using a series of cuts on various detector outputs. For pions, we require a positive status of drift chambers (DC) and scintillator counters (SC) local signals, in addition to the global status reconstructed with the time-based tracking correction. For electron identification, we use good signals from 
all detectors, DC, Cherenkov Counters (CC), SC and the Electromagnetic 
Calorimeter (EC).

Once all particles are identified, we extract our two observables, multiplicity 
ratios and the transverse momentum broadening. Their derivation method is presented in the section \ref{sec:obs}, while a list of our corrections is summarized in the section \ref{sec:corrections}. The budget of our systematical uncertainties is detailed in the section \ref{sec:TotSys}, and the final analysis results are presented and discussed in the next chapter.

@@

First, we apply a fiducial cut on the EC local coordinates to remove electrons detected close to the calorimeter's edges ($U_{EC}>40$~cm, $V_{EC}<360$~cm and 
$W_{EC}<395$~cm). These hits are problematic because part of the electromagnetic shower might be generated outside of the detector. That leads to a wrong measurement of the energy deposited.

To reject pions, we apply a momentum-dependent cut of the energy deposited in the inner ($E_{in}$) and the outer parts ($E_{out})$ of the calorimeter as,

@@

where $\mu = 0.271$ is the mean of the total energy fraction deposited in the 
calorimeter, the parameter $a$ is set to 0.5, and the parameter 
$b$ varies with momenta, as shown in table \ref{tab:ecoutin-par}. The derivation of $b$ was motivated by the non-linear momentum dependence of the energy 
deposited on both EC regions (see figure \ref{eleEC}). As pions are 
minimum ionizing particles, they are expected to loose a constant energy in 
the inner part of the calorimeter (around 30~MeV), regardless of their 
momenta. Therefore, by requesting more than 60~MeV to be deposited, we 
efficiently cut the pion contamination (see figure~\ref{eleECi}).

@@

\begin{figure}[p]
\centering
\includegraphics[width=14cm] {chap5-fig/fig02.jpg} 
\caption {Energy deposited in the inner part of the EC ($E_{in}$) as a function of the energy deposited in its outer part ($E_{out}$) normalized by tracks' momenta. Each panel is for a different momentum range, and the drawn red lines 
illustrate the cuts from equation \ref{einout}.}
\label{eleEC}
\end{figure}

\begin{figure}[tbp]
\centering
\includegraphics[width=9cm] {chap5-fig/fig03.png} 
\caption {Energy deposited in the inner part of the EC as a function of the 
energy deposited in its outer part. The red line illustrates the applied cut.}
\label{eleECi}
\end{figure}

In the CC, the mean number of photo-electrons\footnote{Photo-electrons are 
electrons produced in the front window of the Photo-Multiplier Tube (PMT) by a single photon.} from a high energy electron is expected to be around 10.
However, hadrons can generate noise due to $\delta$ electrons produced in 
the materials of the detector. This signal is expected to be around 
one to two photo-electrons. Hence, to get rid of $\delta$-electrons, we kept only tracks with more than 2.5~photo-electrons, as depicted in figure \ref{delta}.

\begin{figure}[tbp]
\centering
\includegraphics[width=8cm] {chap5-fig/fig04.png} 
\caption {Number of photo-electrons ($\times 10$) per tracks. The red line 
shows the applied cut to remove any low-energy pions' contamination.}
\label{delta}
\end{figure}

To select electrons, we also require that particles are negatively charged\footnote{In the CLAS reconstruction software, the electric charge is determined based on the tracks' bending direction in drift chambers.}.
Positively charged particles verifying electrons' cuts are identified as 
positrons. In the rest of this analysis, we used only events with a single electron and no positron to avoid any confusion between the scattered electron and electrons originating from hadronic decays or a photon conversion.

@@

To identify negatively charged pions, we select negative tracks which didn't verify electrons' identification cuts. Pions are detected in the angular range from $\sim$10 to $\sim$140 degrees using only DC and SC signals.

This identification consists mostly of a time of flight (TOF) test. We define 
the relativistic particles' velocity difference as
$\Delta \beta = \beta_{measured} - {p \over \sqrt{p^2 + m_\pi^2}}$, which is required to be zero within $\pm 0.03$. Figure \ref{PionTOF} illustrates the effect of this cut. As we noticed there are not much negative kaons or anti-protons, we concluded that their contamination must be small (see section~\ref{SysId} for more detailed study).

\begin{figure}[tbp]
\centering
\includegraphics[width=8cm] {chap5-fig/fig05.png} 
\caption {$\Delta \beta$ of negatively charged particles as a function of momentum (GeV/c). The red lines shows the applied cuts to select negative pions.}
\label{PionTOF}
\end{figure}

In principle, a pion identification could be improved using the Cherenkov counter 
for momenta higher than 2.5~GeV/c. But, the observed low CC efficiency (see figure 
\ref{PionCC}) especially at momenta close to the CC threshold ($\sim$25\% at 
2.5~GeV/c and $\sim$50\% at 3~GeV/c) makes its use less compelling. 
Moreover, as only a very small amount of $K^-$s and no ${\bar p}s$ are present on figure \ref{PionTOF}, we decided to not use CC on this pions' identification.

\begin{figure}[tbp]
\centering
\includegraphics[width=8cm] {chap5-fig/fig06.png} 
\caption {Histograms of momentum (GeV/c) for $\pi^-$s. The black, red, and blue distributions are, respectively, for all identified $\pi^-$s, pions that fire Cherenkov counters, and pions that fire CC with more than one photo-electron.}
\label{PionCC}
\end{figure}

@@

The identification of positively charged pions is similar to that of negative 
pions. However, the time of flight plot is significantly more busy (see figure 
\ref{PipTOF}), indicating a significant contamination from $K^+$s and protons at a
higher momentum. As the CC is not efficient enough for a hadron separation, the 
numerous kaons and protons should be removed by the following tighter TOF cuts:

@@

These cuts permit to minimize the kaon contamination, below 2.5~GeV/c, and the 
proton contamination for all momenta. The kaon contribution cannot be avoided, however, it should remain small ($\sim$ 3\% according to simulation) 
with only a small effect on our final results (see section \ref{SysId} 
for more details). Because protons could lead to even more contamination than kaons, 
a stricter cut is used at a high momentum. This cut is also justified by 
HERMES data \cite{Airapetian:2007vu}, which showed a very different behavior of protons compared to other hadrons.

\begin{figure}[tbp]
\centering
\includegraphics[width=8cm] {chap5-fig/pip_data.png} 
\caption {$\Delta \beta$ of positively charged particles as a function of momentum (GeV/c). The red lines shows the applied cuts to select positive pions. The blue lines indicate other particles' theoretical positions, and present from top to 
bottom positrons, muons, kaons and protons.}
\label{PipTOF}
\end{figure}

@@

To differentiate between the two targets and remove the background, we need to 
determine the origin of our final-state particles. As illustrated in figure~\ref{vertex}, the small misalignment of the beam with the CLAS detector led to an artificial shift on different sectors' z-vertex positions. To correct from this effect, we added the shift shown in table~\ref{tab:vertex} when determining each particle's vertex position. Those values were obtained from the fit of the solid target distribution on each sector.

\begin{table}[p]
  \centering
  \begin{tabular}{@{} cc @{}}
    \hline
    Sector & Shift (cm) \\ 
    \hline
    1 & + 0.1 \\ 
    2 & - 0.4 \\ 
    3 & - 0.6 \\ 
    4 & - 0.1 \\ 
    5 & + 0.4 \\ 
    6 & + 0.6 \\ 
    \hline
  \end{tabular}
  \caption{Values used to correct the vertex position on each sector}
  \label{tab:vertex}
\end{table}

The position of our targets could also vary from a run to another, but this did not happen too often in this experiment. Indeed, except for aluminum, other targets remained in the same position within one or two millimeters. All targets' positions, in CLAS coordinates, are given in table \ref{tab:targets}.

\begin{table}[p]
  \centering
  \begin{tabular}{|c|c|c|c|c|c|c|}
    \hline
    Target & Carbon & Al (1) & Al (2) & Iron   & Tin    & Lead   \\ 
    \hline \hline
    Liquid & -30.1  & n/a    & n/a    & -30.2  & -30.1  & -30.1  \\ 
    Solid  & -24.7  & -25.0  & -23.8  & -24.9  & -23.8  & -24.9  \\
    \hline
  \end{tabular}
  \caption{The measured mean vertex position of our nuclear targets relative to the center of CLAS (in cm). Must note that aluminum data are separated in two sets because this target's position has brutally changed. Moreover, during these runs the deuterium target was empty.}
  \label{tab:targets}
\end{table}

\begin{figure}[p]
\centering
\includegraphics[width=10cm] {chap5-fig/Vertex_el_data.png}
\caption {Sector-by-sector electrons' vertex distributions. The latter are reconstructed along the beam direction and relative to the center of CLAS located at 0~cm. The red lines show the applied cuts to select the two targets.}
\label{vertex}
\end{figure}

The detected electrons are associated with the solid target if their vertex 
position is less than 1.5~cm ($\sim$3~$\sigma$) from the values shown on table 
\ref{tab:targets}. For the liquid target, the cut is a bit larger, 2 cm, in order to count for the cryogenic target's size (see figure \ref{vertex}). The pions' vertex was checked against the electron one and was imposed to satisfy this additional condition $| Vz^{e^-} - Vz^{\pi} | < 3$ cm (see figure \ref{fig:dvzpi}).

\begin{figure}[tbp]
\centering
\includegraphics[width=8cm] {chap5-fig/Vertex_pi_data.png}
\caption {The difference between electrons and pions z-vertex positions. $\pi^+$s and $\pi^-$ are plotted, respectively, in the top and bottom panels. The red lines show the applied cuts to select both pions.}
\label{fig:dvzpi}
\end{figure}

@@

To check the quality of the runs\footnote{Runs were formed roughly from two hours
of data, but they were much smaller in case of problems.}, we monitored the ratio of scattered electrons yields obtained from solid and liquid targets. In case of the beam hitting other materials in the beam-line or any detector is not working 
properly, this ratio will be off and indicates problematic runs. A figure 
\ref{DataQ} shows electrons yields' ratios for different nuclear target; we fitted these ratios to extract the mean value for each target, then we discarded runs that were 5~$\sigma$ away from the appropriate mean. Must note that these ratios were consistent with our target's thicknesses given in \cite{Hakobyan:2008zz} except for a carbon target. For that reason, the density of the latter was recently remeasured, and its new value was finally matching our data\footnote{i.e. (1.747+/-0.0007)g/cm$^3$ instead of 2.235 g/cm$^3$}.

\begin{figure}[tbp]
\centering
\includegraphics[width=15cm] {chap5-fig/TargetElRatio.png}
\caption {Ratio of scattered electrons yields (solid over deuterium) for each run.}
\label{DataQ}
\end{figure}

@@

Since we are interested in deep inelastic scattering, we use the following 
cuts: $Q^2 > 1$~GeV$^2$/c$^2$ and $W > 2$~GeV. We also apply a cut on the energy transferred fraction, $y = \nu/E_{beam} < 0.85$. This cut aims to remove the low energy electrons that are significantly affected by radiative effects and sat at the limit of our trigger threshold.

@@

Once the identification is done, the calculation of our two observables and their statistical uncertainties is straightforward. The latter are given by:
\begin{equation}
{\delta \left ( R_A^h \right ) \over R_A^h} = \sqrt{ 1 /N_A^h + 1/ N_A^e +1/N_D^h + 1/N_D^e}
\end{equation}
and
\begin{equation}
\left ( \delta \left ( \Delta \langle P_\perp^2 \rangle \right ) \right )^2 = 
   \left ({\langle P_\perp^4 \rangle - \langle P_\perp^2 \rangle ^2}\right )_A / N_A^h
 + \left ({\langle P_\perp^4 \rangle - \langle P_\perp^2 \rangle ^2}\right )_D / N_D^h.
\end{equation}

However, the implementation of acceptance and radiative corrections is done through some given weights to each event based on its kinematics. The multiplicity ratio then becomes
\begin{equation}
R_A^h (Q^2,\nu,z_h,P_\perp^2) = {{\sum \omega_A^h (Q^2,\nu,z_h,P_\perp^2) / \sum \omega_A^e (Q^2,\nu)} 
                       \over {\sum \omega_D^h (Q^2,\nu,z_h,P_\perp^2) / \sum \omega_D^e (Q^2,\nu)}},
\end{equation}
where $\omega$ is the weight associated with the event, and the sums run over all measured particles. The expression of the transverse momentum broadening remains as
\begin{equation}
\Delta \langle P_\perp^2 \rangle = \langle P_\perp^2 \rangle_A - \langle P_\perp^2 \rangle_D,
\end{equation}
but with
\begin{equation}
\langle P_\perp^2 \rangle = {\sum \left ( P_\perp^2 \times \omega (Q^2,\nu,z_h,P_\perp^2) \right ) \over \sum \omega (Q^2,\nu,z_h,P_\perp^2)}.
\end{equation}

Thus, new expressions of the weighted statistical uncertainties are: 
\begin{equation}
{\delta \left ( R_A^h \right ) \over R_A^h} = 
      \sqrt{ \left ( {\sum {\omega^h_A}^2 \over \left (\sum \omega^h_A \right )^2} \right ) 
           + \left ( {\sum {\omega^e_A}^2 \over \left (\sum \omega^e_A \right )^2} \right ) 
           + \left ( {\sum {\omega^h_D}^2 \over \left (\sum \omega^h_D \right )^2} \right ) 
           + \left ( {\sum {\omega^e_D}^2 \over \left (\sum \omega^e_D \right )^2} \right ) }
\end{equation}
 and 
\begin{equation}
\begin{split}
\left ( \delta \left ( \Delta \langle P_\perp^2 \rangle \right ) \right )^2 = 
   \left ({{\sum \omega^h_A P_\perp^4 }\over{\sum \omega^h_A}} - \left ({\sum \omega^h_A P_\perp^2 }\over{\sum \omega^h_A}\right )^2 \right ) 
         \times \left ( {\sum \left ( {\omega^h_A}^2 \right ) \over \left ( \sum \omega^h_A \right ) ^2 }\right ) \\
 + \left ({{\sum \omega^h_D P_\perp^4 }\over{\sum \omega^h_D}} - \left ({\sum \omega^h_D P_\perp^2 }\over{\sum \omega^h_D}\right )^2 \right ) 
	 \times \left ( {\sum \left ( {\omega^h_D}^2 \right ) \over \left ( \sum \omega^h_D \right ) ^2 }\right ).
\end{split}
\end{equation}

@@

To provide some insight into the quality of our data, we present in figure~\ref{fig:prelim} few preliminary results before the application of any corrections. These preliminary results will be used also to illustrate the effects of the corrections discussed below.

\begin{figure}[htb]
\centering
\includegraphics[width=7.4cm] {chap5-fig/a_RvA.png} 
\includegraphics[width=7.4cm] {chap5-fig/a_RvZ.png} 
\includegraphics[width=7.4cm] {chap5-fig/a_PvA.png} 
\includegraphics[width=7.4cm] {chap5-fig/a_PvNu.png} 
\caption {Results for multiplicity ratios (top) and the transverse momentum 
broadening (bottom) before applying any corrections.}
\label{fig:prelim}
\end{figure}

@@

The acceptance correction consists of applying weights to the experimental data
event by event to correct for any detection inefficiencies of used detectors.
Incidentally, it also corrects for other small detection issues, such as misidentification and re-scattering on detectors' materials. The quality 
of the correction depends on the ability of the simulation to reproduce the 
experiment, the accumulated statistics, and the size of interfering effects, such 
as the bin migration. For this correction, we applied the method from the 
approved and published analysis from the same eg2 data~\cite{ElFassi:2008}.

@@

To correct for acceptance effects, we simulated a total of 100 million events 
per target ($^2$H, C, Fe and Pb) using the PYTHIA \cite{Sjostrand:2006za} 
event generator that was slightly modified to include Fermi motion effects. The 
generated events are processed by CLAS softwares (GSIM, GPP and user\_ana) 
to simulate the detector and the reconstruction process similarly to the experimental data.
% TODO GPP factors
